---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

I'm going to use unzip function then read.csv function. 

```{r, echo = TRUE}

unzip("activity.zip")
activity <- read.csv("activity.csv", stringsAsFactors = F)

activity$date <- as.Date(activity$date)

```

## What is mean total number of steps taken per day?

```{r, echo = TRUE}

Daily_steps <- aggregate(steps ~ date, data = activity, sum)

# histogram of the daily steps 
hist(Daily_steps$steps, xlab = "Steps per day", main = "Histogram of steps per day", breaks = 20, col = "#A40B92")

# mean steps per day 
mean(Daily_steps$steps)

# median steps per day 
median(Daily_steps$steps)
```

## What is the average daily activity pattern?

```{r, echo = TRUE}


interval_average <- aggregate(steps ~ interval, data = activity, mean)

plot(interval_average, type = "l", main = "Average steps per interval", xlab = "Interval (mins)")

max(interval_average$steps)
row_index <- which(interval_average$steps == max(interval_average$steps))

# interval that has the maximum number of steps 
interval_average$interval[row_index]

```


## Imputing missing values

```{r}

## how many NAs 
sum(is.na(activity$steps))

## imputing steps using mean 
steps_imputed <- activity
steps_imputed$steps[is.na(steps_imputed$steps)] <- mean(activity$steps,na.rm =T)

sum(is.na(steps_imputed$steps))

Daily_steps2 <- aggregate(steps ~ date, data = steps_imputed, sum)

# histogram of the daily steps 
hist(Daily_steps2$steps, xlab = "Steps per day", main = "Histogram of steps per day", breaks = 20, col = "#A40B92")

# mean steps per day 
mean(Daily_steps2$steps)

# median steps per day 
median(Daily_steps2$steps)

```

After imputation, mean steps per day stays the same, but median increases slightly to `r format(median(Daily_steps2$steps), scientific = F)`.

## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}

steps_imputed$weekdays <- weekdays(steps_imputed$date)

steps_imputed$weekdays <- sub("Monday|Tuesday|Wednesday|Thursday|Friday", "weekday", steps_imputed$weekdays)

steps_imputed$weekdays <- sub("Saturday|Sunday", "weekend", steps_imputed$weekdays)

steps_imputed$weekdays <- as.factor(steps_imputed$weekdays)

### 

interval_average_weekdays <- aggregate(steps ~ interval*weekdays, data = steps_imputed, mean)

ggplot(interval_average_weekdays, aes(x = interval, y = steps, group = weekdays)) + geom_line() + facet_wrap(~weekdays)


```

